Application of Turchin's method of statistical regularization

### 2018-03-09 AYSS-2017/article/ayss2017-Zeleniy.tex
fixed error
add authors

\title{Application of Turchin's method of statistical regularization}
\author{\firstname{Mikhail} \lastname{Zelenyi}\inst{1,2}\fnsep\thanks{\email{mihail.zelenyy@phystech.edu}} \and
    \firstname{Maria} \lastname{Polyakova}\inst{1,2} 
    }
%
\institute{Institute for Nuclear Research RAS
    \and
    Moscow Institute of Physics and Technology
}
\abstract{%
    During analysis of experimental data, one usually needs to restore a signal after it has been convoluted with some kind of apparatus function. According to Hadamard's definition this problem is ill-posed and requires regularization to provide sensible results. In this article we describe an implementation of the Turchin's method of statistical regularization based on the Bayesian approach to the regularization strategy.
}
%
\maketitle

@@

The resulting prior probability density for the Gaussian random process:
\begin{equation}
  P(\vec{\varphi}|\alpha) = \frac{\det(\alpha\Omega)^{1/2}}{(2\pi)^{N/2}} 
  \exp\left(-\frac{1}{2}\langle\vec{\varphi},\alpha\Omega\vec{\varphi}\rangle\right),
  \label{eq-prior}
\end{equation}
where $\alpha = \alpha(\omega)$~--- parameter of smoothness. In result, we get $\alpha$-depend solution $\hat{S}_{\alpha}[f]$ and now we have problem of picking $\alpha$. In rare cases we can select $\alpha$ manually using known empirical smoothness, but we proposal another way. Information about its value almost never available so we could consider it random variable as well. This yields us hierarchical Bayes model $P(f,\varphi,\alpha) = P(f|\varphi)P(\varphi,\alpha)P(\alpha)$.
Since we don't have any information about $\alpha$ we chose hyperprior $P(\alpha)$ to be flat. Now we have several options to pursue. In general case, we can use fully Bayesian approach  and compute solution to marginalize by $\alpha$ and to use eq. ~\ref{eq:opt}:
    \begin{equation}\label{eq:varAposteriorAlpha}
    P(\varphi) = \int P(\varphi|\alpha) P(\alpha)~d\alpha 
    \end{equation}
or,  to marginalize by $\varphi$ and to integrate by $P(\alpha|f)$:
    \begin{equation}
    P(\alpha|f) \sim \int d\varphi~P(f|\varphi)P(\varphi|\alpha)P(\alpha)
    \qquad\mbox{and}\qquad
    \hat{S}[f] =\frac{1}{Norm} \int d\alpha \hat{S}_{\alpha}[f] P(\alpha|f)
    \label{eq:solveAposteriorAlpha}
    \end{equation}
    using MCMC for sampling $\varphi$ and $\alpha$ (for Gaussian errors it could be reduced to numerical integral). In private case this options can give different solutions if we will use some approximations. Also if distribution $P(\alpha|f)$ is narrow, we can pick $\alpha$ corresponding to most probable $\alpha_0$ and use $P(\varphi|\alpha_0)$ as prior. This is called empirical Bayes. For our problem it produces almost identical results.

>>>>

\title{Application of Turchin's method of statistical regularization}
\author{\firstname{Mikhail} \lastname{Zelenyi}\inst{1,2}\fnsep\thanks{\email{mihail.zelenyy@phystech.edu}} \and
    \firstname{Maria} \lastname{Polyakova}\inst{1,2} \and \firstname{Alexander} \lastname{Nozik}\inst{1,2}\fnsep\thanks{\email{altavir@gmail.com}} \and \firstname{Alexey} \lastname{Khudyakov}\inst{1}\fnsep\thanks{\email{alexey.skladnoy@gmail.com}}  
    }
%
\institute{Institute for Nuclear Research RAS
    \and
    Moscow Institute of Physics and Technology
}
\abstract{%
    During analysis of experimental data, one usually needs to restore a signal after it has been convoluted with some kind of apparatus function. According to Hadamard's definition this problem is ill-posed and requires regularization to provide sensible results. In this article we describe an implementation of the Turchin's method of statistical regularization based on the Bayesian approach to the regularization strategy.
}
%
\maketitle

@@

The resulting prior probability density for the Gaussian random process:
\begin{equation}
  P(\vec{\varphi}|\alpha) = \frac{\det(\alpha\Omega)^{1/2}}{(2\pi)^{N/2}} 
  \exp\left(-\frac{1}{2}\langle\vec{\varphi},\alpha\Omega\vec{\varphi}\rangle\right),
  \label{eq-prior}
\end{equation}
where $\alpha = \alpha(\omega)$~--- parameter of smoothness. In result, we get $\alpha$-depend solution $\hat{S}_{\alpha}[f]$ and now we have problem of picking $\alpha$. In rare cases we can select $\alpha$ manually using known empirical smoothness, but we proposal another way. Information about its value almost never available so we could consider it random variable as well. This yields us hierarchical Bayes model $P(f,\varphi,\alpha) = P(f|\varphi)P(\varphi,\alpha)P(\alpha)$.
Since we don't have any information about $\alpha$ we chose hyperprior $P(\alpha)$ to be flat. Now we have several options to pursue. In general case, we can use fully Bayesian approach  and compute solution to marginalize by $\alpha$ and to use eq. ~\ref{eq:opt}:
    \begin{equation}\label{eq:varAposteriorAlpha}
    P(\varphi) = \int P(\varphi|\alpha) P(\alpha)~d\alpha 
    \end{equation}
or,  to marginalize by $\varphi$ and to integrate by $P(\alpha|f)$:
    \begin{equation}
    P(\alpha|f) \sim \int d\varphi~P(f|\varphi)P(\varphi|\alpha)P(\alpha)
    \qquad\mbox{and}\qquad
    \hat{S}[f] =\frac{1}{Norm} \int d\alpha \hat{S}_{\alpha}[f] P(\alpha|f)
    \label{eq:solveAposteriorAlpha}
    \end{equation}
    using MCMC for sampling $\varphi$ and $\alpha$ (for Gaussian errors it could be reduced to numerical integral). In particular case we can simplify calculations by using following approximations. Also if distribution $P(\alpha|f)$ is narrow, we can pick $\alpha$ corresponding to most probable $\alpha_0$ and use $P(\varphi|\alpha_0)$ as prior. This is called empirical Bayes. For our problem it produces almost identical results.
